---
title: "Week 9 Shiny"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Intro to Shiny
- Shiny is an R package that makes it easy to build interactive web apps straight from R
- A web app updates based on user input/interaction and is made up of UI & server
- The client contains the user interface, that is, buttons and selectors and text boxes and other things that the user can interact with. 
- The server (or backend) is where computation happens, including things like manipulating data and running models.

1. Load `shiny`
2. Create UI with a HTML function
3. Define custom function to create the server
4. Run the app
```{r}
library(shiny)
ui <- fluidPage(
  "Hello world!!!"
)
server <- function(input, output, session) {
  
}
shinyApp(ui = ui, server = server)
```

Ask a question
```{r}
ui <- fluidPage(
  textInput("name", "Enter a name:"),
  textOutput("q")
)
server <- function(input, output) {
  output$q <- renderText({
    paste("Do you prefer cats or dogs,", input$name, "?")
  })
}
shinyApp(ui = ui, server = server)
```

Example
```{r}
# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)

# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}

# Create Shiny app ----
shinyApp(ui = ui, server = server)
```

```{r}
runExample("01_hello") # runs above
runApp("App-1") # runs app
runApp("App-1", display.mode = "showcase") # showcase mode
```

- shiny function	HTML5 equivalent	creates: 
`p`, `h1`, `h6`, `a`, `br`, `div`,	`span`,	`pre`, `code`, `img`, `strong`, `em`, `HTML`
- `img(src = "my_image.png", height = 72, width = 72)`
  - The file must be in a folder named www in the same directory script.
```{r}
ui <- fluidPage(
  titlePanel("My Shiny App"),
  sidebarLayout(
    sidebarPanel(
      h2("Installation"),
      p("Shiny is available on CRAN, so you can install it in the usual way from your R console:"),
      code('install.packages("shiny")'),
      br(),
      br(),
      br(),
      br(),
      img(src = "rstudio.png", height = 70, width = 200),
      br(),
      "Shiny is a product of ", 
      span("RStudio", style = "color:blue")
    ),
    mainPanel(
      h1("Introducing Shiny"),
      p("Shiny is a new package from RStudio that makes it ", 
        em("incredibly easy "), 
        "to build interactive web applications with R."),
      br(),
      p("For an introduction and live examples, visit the ",
        a("Shiny homepage.", 
          href = "http://shiny.rstudio.com")),
      br(),
      h2("Features"),
      p("- Build useful web applications with only a few lines of code—no JavaScript required."),
      p("- Shiny applications are automatically 'live' in the same way that ", 
        strong("spreadsheets"),
        " are live. Outputs change instantly as users modify inputs, without requiring a reload of the browser.")
    )
  )
)

server <- function(input, output) {
  
}

shinyApp(ui = ui, server = server)
```


Shiny widgets
`function`	widget
`actionButton`	Action Button
`checkboxGroupInput`	A group of check boxes
`checkboxInput`	A single check box
`dateInput`	A calendar to aid date selection
`dateRangeInput`	A pair of calendars for selecting a date range
`fileInput`	A file upload control wizard
`helpText`	Help text that can be added to an input form
`numericInput`	A field to enter numbers
`radioButtons`	A set of radio buttons
`selectInput`	A box with choices to select from
`sliderInput`	A slider bar
`submitButton`	A submit button
`textInput`	A field to enter text

```{r}
ui <- fluidPage(
  titlePanel("censusVis"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Create demographic maps with 
               information from the 2010 US Census."),
      
      selectInput("var", 
                  label = "Choose a variable to display",
                  choices = list("Percent White", 
                                 "Percent Black",
                                 "Percent Hispanic", 
                                 "Percent Asian"),
                  selected = "Percent White"),
      
      sliderInput("range", 
                  label = "Range of interest:",
                  min = 0, max = 100, value = c(0, 100))
    ),
    
    mainPanel()
  )
)
```


`actionButton("action", label = "Action")`
https://shiny.rstudio.com/tutorial/written-tutorial/lesson3/

- Hazard function
- Royston-Parmar model
- tpt: thinnest point thickness
- period: follow up time 
- status: whether or not they had CXL


```
setwd("/home/jovena/internship/RShinyCXL/shinyapp")
library(rsconnect)
rsconnect::deployApp()
```

- setwd("/home/jovena/internship/summer_internship/Week-9")

















